Towards understanding the central limit theorem by learning Python basics

### 2021-02-18 paper/article.tex
Add information on the first two sheets

% Own definitions (Zoltan)
\usepackage{listingsutf8,newtxtt,xcolor,graphicx,xurl}
\lstset{language=Python,
  basicstyle=\ttfamily,
  keywordstyle=\bfseries\color{magenta},
  stringstyle=\color{red},
  numberstyle=\color{blue}
  breaklines=true,
  showstringspaces=false,
  backgroundcolor=\color{yellow!30},
  ndkeywordstyle=\color{yellow},
  commentstyle=\color{green},
  identifierstyle=\color{black},
  literate=
  {á}{{\'a}}1 {é}{{\'e}}1 {í}{{\'i}}1 {ó}{{\'o}}1 {ú}{{\'u}}1
  {Á}{{\'A}}1 {É}{{\'E}}1 {Í}{{\'I}}1 {Ó}{{\'O}}1 {Ú}{{\'U}}1
  {à}{{\`a}}1 {è}{{\`e}}1 {ì}{{\`i}}1 {ò}{{\`o}}1 {ù}{{\`u}}1
  {À}{{\`A}}1 {È}{{\'E}}1 {Ì}{{\`I}}1 {Ò}{{\`O}}1 {Ù}{{\`U}}1
  {ä}{{\"a}}1 {ë}{{\"e}}1 {ï}{{\"i}}1 {ö}{{\"o}}1 {ü}{{\"u}}1
  {Ä}{{\"A}}1 {Ë}{{\"E}}1 {Ï}{{\"I}}1 {Ö}{{\"O}}1 {Ü}{{\"U}}1
  {â}{{\^a}}1 {ê}{{\^e}}1 {î}{{\^i}}1 {ô}{{\^o}}1 {û}{{\^u}}1
  {Â}{{\^A}}1 {Ê}{{\^E}}1 {Î}{{\^I}}1 {Ô}{{\^O}}1 {Û}{{\^U}}1
  {Ã}{{\~A}}1 {ã}{{\~a}}1 {Õ}{{\~O}}1 {õ}{{\~o}}1
  {œ}{{\oe}}1 {Œ}{{\OE}}1 {æ}{{\ae}}1 {Æ}{{\AE}}1 {ß}{{\ss}}1
  {ű}{{\H{u}}}1 {Ű}{{\H{U}}}1 {ő}{{\H{o}}}1 {Ő}{{\H{O}}}1
  {ç}{{\c c}}1 {Ç}{{\c C}}1 {ø}{{\o}}1 {å}{{\r a}}1 {Å}{{\r A}}1
  {€}{{\euro}}1 {£}{{\pounds}}1 {«}{{\guillemotleft}}1
  {»}{{\guillemotright}}1 {ñ}{{\~n}}1 {Ñ}{{\~N}}1 {¿}{{?`}}1,
  columns=fullflexible,
  keepspaces=true
}
\makeatletter
\def\lst@outputspace{{\ifx\lst@bkgcolor\empty\color{white}\else\lst@bkgcolor\fi\lst@visiblespace}}
\makeatother

@@

In this paper we demonstrate our proposal by explaining an email based course, sent out
to 4 prospective mathematics teachers, allowing them an arbitrary time to solve the problems being set.
The students claimed that they did not have any deeper knowledge in programming formerly, but all of
them already studied probability theory at university level. Their knowledge, however, was
not yet checked via examination. The course consisted of 7 sets of homework assignments
and took place between November 2020 and January 2021, at the Johannes Kepler University of Linz, Austria.

\section{Preparations}

@@

The first example is shown in Fig.~\ref{main-py-window}. We kept the original version to highlight
the possibility to write special characters in the Python code (we used Python version 3),
including accented German letters. In the first set of problems each student had to try
the code on his or her own on the website \url{https://repl.it/languages/python3},
by paying attention on the correct indentation as well---the students, however, did not have
to type the code but just copy it from the email into the programming window.
We also instructed all students with an
explanatory screenshot how the program should be started and the output obtained.

@@

In this introductory set the first question was to make the program work by running it.
To give a somewhat creative assignment we also asked how the output 3 can be changed to
the number 6. The outcome of this puzzle will be analyzed later in Section \ref{problem-sets}.

In fact, we created a couple of other programs in advance, but always waited for
all solutions turned in by the students before finalizing the next set of problems.
We \textit{did not} give any feedback on their solutions. Instead, if we felt that
some difficulties arose, then we tried to cover them in the next set of problems.
The reason behind this was to test if it is possible to create a set of problems
\textit{in advance} for a possible \textit{static} learning path. Even if our experiment
was not fully static, for a future project now we already have a static set of materials.

@@

\section{The problem sets}
\label{problem-sets}

>>>>

% Own definitions (Zoltan)
\usepackage{listingsutf8,newtxtt,xcolor,graphicx,xurl}
\lstset{language=Python,
  numbers=left,
  basicstyle=\ttfamily,
  keywordstyle=\bfseries\color{magenta},
  stringstyle=\color{red},
  numberstyle=\color{blue},
  breaklines=true,
  showstringspaces=false,
  backgroundcolor=\color{yellow!30},
  ndkeywordstyle=\color{magenta!70},
  commentstyle=\color{green},
  identifierstyle=\color{black},
  literate=
  {á}{{\'a}}1 {é}{{\'e}}1 {í}{{\'i}}1 {ó}{{\'o}}1 {ú}{{\'u}}1
  {Á}{{\'A}}1 {É}{{\'E}}1 {Í}{{\'I}}1 {Ó}{{\'O}}1 {Ú}{{\'U}}1
  {à}{{\`a}}1 {è}{{\`e}}1 {ì}{{\`i}}1 {ò}{{\`o}}1 {ù}{{\`u}}1
  {À}{{\`A}}1 {È}{{\'E}}1 {Ì}{{\`I}}1 {Ò}{{\`O}}1 {Ù}{{\`U}}1
  {ä}{{\"a}}1 {ë}{{\"e}}1 {ï}{{\"i}}1 {ö}{{\"o}}1 {ü}{{\"u}}1
  {Ä}{{\"A}}1 {Ë}{{\"E}}1 {Ï}{{\"I}}1 {Ö}{{\"O}}1 {Ü}{{\"U}}1
  {â}{{\^a}}1 {ê}{{\^e}}1 {î}{{\^i}}1 {ô}{{\^o}}1 {û}{{\^u}}1
  {Â}{{\^A}}1 {Ê}{{\^E}}1 {Î}{{\^I}}1 {Ô}{{\^O}}1 {Û}{{\^U}}1
  {Ã}{{\~A}}1 {ã}{{\~a}}1 {Õ}{{\~O}}1 {õ}{{\~o}}1
  {œ}{{\oe}}1 {Œ}{{\OE}}1 {æ}{{\ae}}1 {Æ}{{\AE}}1 {ß}{{\ss}}1
  {ű}{{\H{u}}}1 {Ű}{{\H{U}}}1 {ő}{{\H{o}}}1 {Ő}{{\H{O}}}1
  {ç}{{\c c}}1 {Ç}{{\c C}}1 {ø}{{\o}}1 {å}{{\r a}}1 {Å}{{\r A}}1
  {€}{{\euro}}1 {£}{{\pounds}}1 {«}{{\guillemotleft}}1
  {»}{{\guillemotright}}1 {ñ}{{\~n}}1 {Ñ}{{\~N}}1 {¿}{{?`}}1,
  columns=fullflexible,
  keepspaces=true
}
\makeatletter
\def\lst@outputspace{{\ifx\lst@bkgcolor\empty\color{white}\else\lst@bkgcolor\fi\lst@visiblespace}}
\makeatother

@@

In this paper we demonstrate our proposal by explaining an email based course, sent out
to 4 prospective mathematics teachers, allowing them an arbitrary time to solve the problems being set.
The students claimed that they did not have any deeper knowledge in programming formerly, but all of
them already studied probability theory at university level. Their knowledge, however, was
not yet checked via examination. The course consisted of 7 sheets of homework assignments
and took place between November 2020 and January 2021, at the Johannes Kepler University of Linz, Austria.

\section{Preparations}
\label{preps}

@@

The first example is shown in Fig.~\ref{main-py-window}. We kept the original version to highlight
the possibility to write special characters in the Python code (we used Python version 3),
including accented German letters. In the first sheet of problems each student had to try
the code on his or her own on the website \url{https://repl.it/languages/python3},
by paying attention on the correct indentation as well---the students, however, did not have
to type the code but just copy it from the email into the programming window.
We also instructed all students with an
explanatory screenshot how the program should be started and the output obtained.

@@

In this introductory sheet the first question was to make the program work by running it,
and then explain its meaning line by line.
To give a somewhat creative assignment we also asked how the output $2/6$ can be changed to
the $3/6$. The outcome of this puzzle will be analyzed later in Section \ref{problem-sheets}.

It was expected that the students solve the sheet problems on their own, without
any communication. However, external help like using Internet was allowed for them.
The students were informed that about 6--8 sheets should be solved in the project,
and the required time for each should not exceed one hour.

In fact, we created a couple of other programs in advance, but always waited for
all solutions turned in by the students before finalizing the next sheet of problems.
We \textit{did not} give any feedback on their solutions. Instead, if we felt that
some difficulties arose, then we tried to cover them in the next sheet of problems.
The reason behind this was to test if it is possible to create a set of problems
\textit{in advance} for a possible \textit{static} learning path. Even if our experiment
was not fully static, for a future project now we already have a static set of materials.

@@

\section{The problem sheets}
\label{problem-sheets}

\subsection*{Sheet 1}
The first sheet has already been mentioned in Section \ref{preps}.
The expected explanation of the Python program was solved well by the students.
All of them managed to write a correct answer to the creative question, namely
\begin{lstlisting}[language=Python,firstnumber=5]
  if Augenzahl > 3:
\end{lstlisting}
One student mentioned that the same output can be obtained with this change:
\begin{lstlisting}[language=Python,firstnumber=2]
günstig = 1
\end{lstlisting}
when line 5 is kept in its original form. Of course, this answer makes sense
only from the programmer's point of view, and it has less to do with probability.

After receiving the results for the first sheet we had the impression that
the students got familiar with the basics of the required technology.

\subsection*{Sheet 2}

The second sheet consisted of the following four questions:

\begin{enumerate}
\item Consider the following program:
\lstinputlisting[language=Python]{../progs/en_US/2.py}

\item Explain its meaning line by line, in one sentence for each.

\item Give the probability space where the possible outcomes are counted.
Which event has been identified here as ``desired''?

\item Modify the program to get a mechanical solution for the following question:

\begin{quote}
One fair die is rolled three times. Compute the probability for the event that the outcomes
correspond to a monotone increasing sequence.
\end{quote}

\end{enumerate}

We asked the students to try to solve the sheet problems without using any external help.
In case of lack of success
they were allowed to use a website that explains how to use if-then-else conditions in Python.

All students faced the quoted question during the introductory course in probability, but that time the solution
was expected to follow a purely pencil-and-paper method.

\subsubsection*{The results}

This sheet was much more difficult than the first one. One student reported that about an hour was required to solve it.
We learned that one student has a misconception by mixing relative frequency and probability---it was
a kind of surprise for this student that the program computes \textit{probability}, even if
there is no randomness built in. (In this case the students' misconception was corrected
by an explanation given in an email answer. Here we made an exception and forced
communication to clarify this basic concept.)

In general we were satisfied with the students' answers. All of them learned how three loops can
be nested to iterate on $\Omega=[1,2,\ldots,6]^3$. Multiple ways were found to express monotonicity,
including both the construct with an \texttt{and} clause and the short form
\texttt{outcome_die_1 <= outcome_die_2 <= outcome_die_3} which is one of Python's strengths.

In our experience it turned out that with little work a definitely large probability space
can be studied. In fact, manual check of $[1,2,\ldots,6]^3$ (or, equivalently, deriving the sum
$1+3+6+10+15+21$) is much more time consuming than writing a suitable program.
